### 2019-02-14 coverletter.tex
Updates from Overleaf

\documentclass[11pt]{article}
\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{fancyhdr} \pagestyle{fancy}
%\renewcommand{\baselinestretch}{1.0} %double linespread or \linespread{1.6}
\setlength{\textwidth}{16cm}
\setlength{\textheight}{23cm}
\setlength{\hoffset}{-1.6cm}
\setlength{\voffset}{-1.cm}

\begin{document} %%

\begin{minipage}[t]{0.55\textwidth}
~ \vspace{0pt}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
Wenjie Lei \\
Department of Geosciences \\
Princeton University \\
Princeton, New Jersey 08540\\
USA \\
~ \\
\today \\ 
\end{minipage} \hspace{0cm}\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\par
\noindent
Ana Ferreira\\
Editor \\
\textit{Geophysical Journal International}\\
~ \\
Dear Editor Ferreira,\\
\par \noindent
We are pleased to submit the original research article entitled ``Global Adjoint Tomography -- Model GLAD-M25" for publication in the \textit{Geophysical Journal International}. 
We would like to suggest the following potential 
reviewers. \\
~ \\
David Al-Attar, University of Cambridge \\
Email: da380@cam.ac.uk\\
~ \\
Heiner Igel, University of M\"unich \\
Email: igel@geophysik.uni-muenchen.de \\
~ \\
Qinya Liu, University of Toronto\\
Email:  liuqy@physics.utoronto.ca \\
~ \\
Jeroen Ritsema, University of Michigan \\
Email:  jritsema@umich.edu \\
~ \\
Carl Tape, University of Alaska Fairbanks \\
Email:  ctape@alaska.edu \\
~ \\
Jeannot Trampert, University of Utrecht \\
Email: J.A.Trampert@uu.nl \\
~ \\

\par  \noindent
Thank you for your consideration.\\
\par \noindent
Sincerely, \\
\par \noindent
Wenjie Lei
\end{document} %%